General purpose gene annotation using deep learning

### 2020-04-07 paper/main.tex
dCNN stuff started

We also implemented a dilated CNN baseline as this kind of architecture is widely used when working with DNA data as input. Details of the neural architecture search are given in Section \ref{sec:dcnn} and results are shown in Table \ref{tab:results}.

The model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously. This grouping enabled us to train effectively with far longer sequences than usual. A class wise evaluation including the calculation of the Genic F1 score is performed on the validation set after every epoch.

@@

\section{Supplementary Material}
\subsection{Dilated CNN Baseline Network Search}
\label{sec:dcnn}

@@

\subsection{Hyperparameters}

\begin{table}[!t]
\processtable{Hyperparameters\label{tab:params}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Value \\
\midrule
Layers & 4\\
Units per layer & 256\\
Learning rate & 1e-4\\
Optimizer & Adam\\
\botrule
\end{tabular}}{}
\end{table}

\subsection{Training Data statistics}
\label{sec:training_data}

\begin{table}[!t]
\processtable{Data group statistics for our training data\label{tab:statistics_training}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\botrule
\end{tabular}}{The description of Table \ref{tab:statistics_all} applies here as well.}
\end{table}

>>>>

We also implemented a dilated CNN (dCNN) baseline as this kind of architecture is widely used when working with DNA data as input. Details of the neural architecture search are given in Section \ref{sec:dcnn}.

Our final model was trained with 10 bases of genomic sequence as input during each time step and produces individual predictions for each of those 10 bases simultaneously. This grouping enabled us to train effectively with far longer sequences than usual. A class wise evaluation including the calculation of the Genic F1 score is performed on the validation set after every epoch.

@@

\section{Supplementary Material}

\subsection{Dilated CNN Baseline Network Search}
\label{sec:dcnn}
\begin{table}[!t]
\processtable{Parameter search space for the dilated CNN baseline\label{tab:dcnn}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Value \\
\midrule
Layers & 4\\
Units per layer & 256\\
Learning rate & 1e-4\\
Optimizer & Adam\\
\botrule
\end{tabular}}{Above are all parameters used during neural architecture search for the dilated CNN baseline. The same overall space was used for plant and animal data. We did, however, run multiple distinct searches that sometimes only operated over a subset of the given parameters. This was done as those seemed to be the most promising. We for example restricted the search space of the number of LSTM layers to the highest 3 values in later runs. Decisions were guided by the Genic F1 on the validation set of exactly the same data we trained our final LSTM models with.}
\end{table}

\begin{table}[!t]
\processtable{Final Parameters of dilated CNN baseline\label{tab:dcnn_best}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Animal Value & Plant Value\\
\midrule
Layers & 4 & \\
Units per layer & 256 & \\
Learning rate & 1e-4 & \\
Optimizer & Adam & \\
\botrule
\end{tabular}}{The parameters of the dilated CNN model that performed the best according to the Genic F1 of the validation data of the respective training genomes.}
\end{table}

@@

\subsection{Hyperparameters}
\begin{table}[!t]
\processtable{Hyperparameters\label{tab:params}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Value \\
\midrule
Layers & 4\\
Units per layer & 256\\
Learning rate & 1e-4\\
Optimizer & Adam\\
\botrule
\end{tabular}}{}
\end{table}

\subsection{Training Data statistics}
\label{sec:training_data}
\begin{table}[!t]
\processtable{Data group statistics for our training data\label{tab:statistics_training}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\botrule
\end{tabular}}{The description of Table \ref{tab:statistics_all} applies here as well.}
\end{table}
